---
title: "Tract Vaccination Priority Rankings"
author: "Sam Powers, Michele Claibourn"
date: "3/18/2021"
#output: html_document
header-includes:
 - \usepackage[default]{sourcesanspro}
 - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
fontsize: 11pt
output: 
  html_document:
    df_print: kable
---

<style>
.leaflet {
    margin: auto;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(extrafont)
library(leaflet)
library(tidycensus)
library(RColorBrewer)
library(knitr)
library(rmarkdown)
library(DT)
loadfonts

load("geo_rankings_indigenous.Rdata")
tract_dimensions <- read_csv("../data/tract_dimensions.csv") %>%
  mutate(GEOID = as.character(GEOID))


```


## This Report
This report presents a composite score to help prioritize census tracts for vaccination according to their alignment with the BRHD vaccination equity strategy priorities. 

## BRHD Census Track Vaccination Priority Score
The [BRHD Vaccination Equity Strategy](https://www.vdh.virginia.gov/content/uploads/sites/196/2021/03/COVID-19-Vaccination-Equity-Strategy-BRHD.pdf) "prioritize[s] vaccination events in neighborhoods and apartment buildings with compounding factors including, at a minimum, low life expectancy, poverty and race and making vaccine available to anyone who lives in that neighborhood." Going beyond that minimum, we include prioritizing populations with medical conditions that increase risk of severe illness from the virus that causes COVID-19. This means we have to balance the following 6 track-level variables when determining vaccination priority:

* % Population Black
* % Population Indigenous American
* % Population Latinx
* % Population in Poverty
* Tract-level Life Expectancy
* % Population with at-risk health outcomes [CDC](https://www.cdc.gov/coronavirus/2019-ncov/need-extra-precautions/people-with-medical-conditions.html)

Our statistical analyses of these six variables suggest it is reasonable to combine them into one composite score. In other words, we created one summary variable that, when orderd, ranks tracts in terms of their priority on the six equity variables. We present that composite score below. 


## Priority Tract Scoring Table 
The following table provides the overall index, scoring tracts by the percent of residents who are Black, who are Hispanic or Latinx, and who are experiencing poverty along with the average life expectancy of residents within the tract and the overall prevalance of health outcomes identified by CDC as risk factors: cancer (except skin cancer), chronic kidney disease, chronic obstructive pulmonary disease, diabetes, obesity, and smoking. 

The indicator values provided in the table are normed via a min-max transformation $$I - \frac{x - min(x)}{max(x) - min(x)}$$. The initial estimated values are provided in a second table below.

``` {r}
rankings_geo_indigenous %>%
  as_tibble() %>%
  select(NAME, Score = overall, `Black Residents` = blackE, `Latinx Residents`= ltnxE, `Indigenous Residents` = indigE, `Poverty` = povrateE, `Life Expectancy` = lifeexpBRHD, `Adverse Health Outcomes` = health_outcomes )  %>%
     mutate(across(where(is.numeric), ~round(.x, 2))) %>%
  datatable()
```


## Priority Tract Scoring Map 

The map visualizes the composite score from the table above to ease identification of census tract locations. With a list of addresses for current vaccination distribution sites, we could add these to the map for easier visualization.

```{r, warning = FALSE, message = FALSE}

mycol <- colorRampPalette(brewer.pal(8, "YlGnBu"))(10)
pal <- colorNumeric(palette = mycol,
                    domain = rankings_geo_indigenous$overall)

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  # addTiles() %>% # to show streets more prominently
  addPolygons(data = rankings_geo_indigenous, 
              fillColor = ~pal(rankings_geo_indigenous$overall), 
              fillOpacity = 0.5, 
              color = "white",
              weight = 2, 
              
              smoothFactor = 0.2, 
              popup = paste0("Priority Score: ",  round(rankings_geo_indigenous$overall, 2), "<br>",
                             rankings_geo_indigenous$NAME, "<br>"
                             ),
              
              highlight = highlightOptions(
                weight = 5,
                fillOpacity = 0.7,
                bringToFront = TRUE)
              ) %>%
  addLegend(pal = pal, 
            values = rankings_geo_indigenous$overall, 
            position = "bottomright", 
            opacity = 0.5,
            title = "Vacination Priority Index")
```


## Composite Indicator Values Table 

The measures used to generate the composite score are provided below for reference.

``` {r, message = FALSE}

tract_dimensions %>%
  left_join(rankings_geo_indigenous %>%
              select(GEOID, overall, NAME)
            )  %>%
  arrange(-overall) %>%
  select(`County Name` = countyname, NAME, overall, everything(), -GEOID, -geometry,
         -whiteE, -asianE, -multiE, -othraceE, -lifeexpCDC, -cpovrateE) %>%
  separate(NAME, c(NA, "Tract", NA), sep = "Census Tract|,") %>%
 rename_with( ~str_to_sentence(str_replace_all(str_replace_all(.x, "_outcome", ""), "_", " ")) , contains("outcome")
        ) %>%
  rename(
    `Priority Index` = overall,
    `Total Pop` = totalpopE,
    `% Black` = blackE,
    `% Indigenous` = indigE,
    `% Latinx` = ltnxE,
    `% Poverty` = povrateE,
    `Life Expectancy` = lifeexpBRHD
  ) %>%
   mutate(across(where(is.numeric), ~round(.x, 2))) %>%
  datatable()
  

```

## Data Sources

* Percent of residents who are Black: American Community Survey 5-year estimates, 2015-2019, Table DP05
* Percent of residents who are Hispanic or Latinx: American Community Survey 5-year estimates, 2015-2019, Table DP05
* Percent of residents experincing poverty: American Community Survey 5-year estimates, 2015-2019, Table S1701
* Average Life Expectancy: Blue Ridge Health District estimates 
* Health prevalence outcomes: [CDC Places estimates](https://chronicdata.cdc.gov/500-Cities-Places/PLACES-Local-Data-for-Better-Health-Census-Tract-D/cwsq-ngmh)








